Cartogram Mapping and Application to Cancer Data.

### 2019-06-19 review_article.Rmd
Choropleth mapping with references

Choropleth mapping methods explore a relationship across a geographical space, utilising administrative or other boundaries. In thematic maps colours, symbols or patterns are used represent levels, or rates, of the information communicated. @CIBMUK believes these 'representations are highly intuitive' as the familiar shapes help those looking at the maps to infer spatial relationships. 
They are true maps, used for visual inspection and can help to direct hypotheses about the statial patterns across the physical space. 
These true maps rely on visual perception, a fallible metric as it can 'weight the shades from the largest areas more heavily.' The inferences derived from visual inspection, and the hypotheses developed can suffer from 'bias'. The use of thematic disease mapping was motivated by a 'desire to evaluate geographical patterns' @DMAHP. This approach has been seen as a 'trigger for questions' which drives research of spatial distributions.  @MACM also recognises the intuition that can be derived from choropleths. However, his critique is that they must be 'validated by rigorous statistical analyses.'

As a cancer mapping tool choropleths are often used. Specifically in epidemiology, the study of the distribution of health-related states and events. Almost 100 years of cancer maps have explored the US and the UK with increased effectiveness in presentation of unbiased rates. @DMAHP cites Cruickshank's (1947) discussion of 'formal statistical assessment of the spatial pattern' as another major advancement.

The data collection methods, and administrative control within regions lends itself to the presentation of cancer mortality rates across regions.
Data is now presented often age adjusted, and as relative rates of risk. Some also present changes in the rates

@@

@@

<!--
## Temporal displays
-->

@@

Plotting centroids on top of geographies. (Size is kept constant)
Hexmaps

@@

@SEMA
'Disease mapping is carried out to summarise spatial and spatio-temporal variation in risk.'
Decriptive statistic of the relationship between geography and a disease.
population, exposure, and health data may have pointdata or aggregated summaries (countdata)
potentially be subject to a number of inaccuracies
diagnostic error or misclassification

@@

Choropleth Maps, relationship between regions and spatial location. Able to represent all the occurences within areas more cohesively, as total counts, or colours.

>>>>

Choropleth mapping methods explore a relationship across a geographical space.
Thematic maps involve drawing administrative or other boundaries and filling the regions with colours, symbols or patterns to present levels, or rates, of the information communicated. @CIBMUK believes these 'representations are highly intuitive' as the familiar shapes help those looking at the maps to visually infer spatial relationships. 
The visualisations are true maps, used for visual inspection and can help to direct hypotheses about the spatial patterns across the physical space. 
These true maps rely on visual perception, a fallible metric as it can 'weight the shades from the largest areas more heavily.' The inferences derived from visual inspection, and the hypotheses developed can suffer from 'bias'. The use of thematic disease mapping was motivated by a 'desire to evaluate geographical patterns' @DMAHP. This technique has been seen as a 'trigger for questions' which drives research of spatial distributions. @MACM also recognise the intuition that can be derived from choropleths with the critique that they must be 'validated by rigorous statistical analyses.'

In epidemiology, choropleths are often used as a mapping tool in the study of the spatial distribution. Almost 100 years of cancer maps have explored the US and the UK with increased effectiveness in presentation of unbiased rates. 
The data collection methods, and administrative control within regions lends itself choropleth visualisation, for the presentation of cancer mortality rates across regions. Mortality and incidence rates are now presented age adjusted, and as relative rates of risk across the population.

@DMAHP cites Cruickshank's (1947) discussion of using visuals as 'formal statistical assessment of the spatial pattern' as a major advancement.
@HEDP considered maps a research tool, dots maps use spots, choropleths and thematic maps shade areas and isopleths use isolines. These  techniques were commonly used before the advent of computer-assisted cartography.

<!--
-Atlases
-methods they use
-->

@@

@HEDP argues that 'cancer occurs in people, not in geographical areas' and the map base should reflect this to avoid allocating 'undue prominence' to rural areas.
geographical and demographic base maps could be used after digital computer methods became available.

@@

Used in displays of the UK by Howe in 1986 cited by @HEDP
<!--
## Temporal displays
-->

Computer generated map examples:  @HEDP
(Hopps et al. 1968; Armstrong 1972). There
has followed a flood of disease atlases, mainly concentrating on the modem problems of cancer and degenerative diseases from countries as scattered as the United States (Burbank 1971; Mason et al. 1975, 1976; Pickle et al. 1987), the Soviet Union (Levin 1980), Japan (Shigematsu 1977), the Federal Republic of Germany

@@

Plotting centroids on top of geographies. (Size is kept constant)
Hex maps

@@

@SEMA
'Disease mapping is carried out to summarise spatial and spatio-temporal variation in risk.'
Descriptive statistic of the relationship between geography and a disease.
population, exposure, and health data may have point data or aggregated summaries (count data)
potentially be subject to a number of inaccuracies
diagnostic error or misclassification

@@

Choropleth Maps, relationship between regions and spatial location. Able to represent all the occurrences within areas more cohesively, as total counts, or colours.
